A molecular dynamics study of the solvation of \ce{CO_2}, water, and some organic compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-07-25 ionic_liquids.tex
strategies of IL revised

%\documentclass[3p,preprint]{elsarticle}
\documentclass[3p,twocolumn]{elsarticle}
\usepackage[draft,hidelinks,colorlinks,citecolor=blue]{hyperref}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{commath}
\usepackage{amssymb}
\usepackage{scalerel}
\usepackage{natbib}

@@

%% Numbered without titles
\bibliographystyle{model1a-num-names}

@@

%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

@@

Recently, the potentiality of ILs for \ce{CO_2} capture has been called into question by Carvalho \textit{et al.} \cite{Carvalho_2016}, who warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposed absorption capability of certain ILs. Although most thermodynamic models are formulated in terms of molar fractions, a key factor that dictates the absorption process efficiency is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent. By expressing the solubility values in terms of molality and performing an exhaustive comparison between various ILs, Ref. \cite{Carvalho_2016} shows that most systems which promote a physical absorption of \ce{CO_2} exhibit a solubility similar or even lower than that observed for the heavy alkane eicosane. This is true for the ``popular'' \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}; the only exception being the family of ILs containing the anion tetracyanoborate \ce{[B(CN)_4]^-}.

@@

In the following, the subscripts ``$s$'' and ``$\text{IL}$'' refer to the solute and the solvent (ionic liquid), respectively. For a system containing $N_{\text{IL}}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (P) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is formally defined as
\begin{equation}
\Delta G_{\text{sim}} \equiv k_b T \ln \frac{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 1)}{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 0)}, 
\end{equation}
where superscript ``$l$'' refers to the liquid solution, $\bigtriangleup$ denotes the isobaric-isothermal partition function, and $k_b$ is the Boltzmann's constant. Shirts \textit{et al.} \cite{Shirts_2003} rigorously derived the working expression that relates $\Delta G_{\text{sim}}$ to what is experimentally known as the solvation free energy. They found out that
\begin{equation}
\begin{split}
\label{eq:free_solv}
 \Delta G_{\text{solv}}& =  k_b T \ln \left( \frac{P_s}{c_s^{\ast} k_b T} \right)\\ &= \Delta G_{\text{sim}} - k_bT \ln \left( \frac{V^{\ast}}{V_1} \right) ,
\end{split}
\end{equation}
where $V^{\ast}$ and $V_1$ refer to the mean volumes of the systems specified by ($N_s^l-1$,$N_{\text{LI}}$,P,T) and ($N_s^l$,$N_{\text{LI}}$,P,T), respectively; $c_s^{\ast}$ = $N_s^l/V_1$ is the numerical concentration of the solute in the liquid phase and $P_s$ is the partial pressure of the solute in the gas phase. It should be noted that in the derivation of Eq.~\ref{eq:free_solv}, the gas phase was treated as an ideal gas.

@@

\subsection*{Henry constant}
If both $P_s$ and $c_s^{\ast}$ were not large, we can resort to Eq.~\ref{eq:free_solv} to predict Henry's law constant $\text{K}_s$, which is defined as \cite{Prausnitz}
\begin{equation}
\text{K}_s =\frac{P_s}{x_s}.
\end{equation}
Therefore, by simply expressing Eq.~\ref{eq:free_solv} in terms of the molar fraction $x_s$ and assuming that MW (molecular weight of the solution) = $\text{MW}_{\scaleto{IL}{4pt}}$, as well as $\rho$ (mass density of the solution) = $\rho_{\scaleto{IL}{4pt}}$, we obtain that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_{\scaleto{IL}{4pt}} R \, T \exp(\beta \, \Delta G_{\text{sim}}),
\end{equation}
where $\tilde{\rho}_{\scaleto{IL}{4pt}}$ is the molar density of the solvent and $R$ is the gas constant.

@@

If one makes $N_s^{l}$ = 1, and substitute Eqs.~\ref{eq:mu_res} and~\ref{eq:mu_res_0} into Eq.~\ref{eq:gamma}, the infinite-dilution activity coefficient, $\gamma^{\infty}_s$, can be predicted according to the following equation:
\begin{equation}
\label{eq:gamma_inf}
\ln \gamma^{\infty}_s = \frac{1}{k_b T} \left( \Delta G_{\text{sim}} - \Delta G_{\text{sim}}^{0} \right) + \frac{c_{\scaleto{LI}{4pt}}^{\, \ast}}{c_{\, s}^{\, \ast,0}},
\end{equation}
where $c_{\scaleto{IL}{4pt}}^{\, \ast}$ y $c_{\, s}^{\, \ast,0}$ are the numerical densities of the solvent and the pure solute, respectively.

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} are found at pressure values for which, in principle, the ideal gas assumption may be not valid. In this case, we shall correct the ideal gas contribution to the chemical potential, as follows
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + RT \ln \phi_s,
\end{equation}
where $\phi_s$ is the fugacity coefficient of the solute. This equation, together with the fact that $f^{g}_s$ = $P_s \phi_s$, are used to evaluate $\mu^{l}_s$ = $\mu^{g}_s$, which yields
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = \exp  \left( \beta \Delta G_{\text{sim}} \right) +  k_b T c_s^{\ast}, 
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase. Having calculated $f^{g}_s$, $P_s$ can be predicted using the following equation
\begin{equation}
\label{eq:fgas}
f^{g}_s = P_s \exp\left[ \frac{B_s(T) P_s}{R T} \right],
\end{equation}
where $B_s$ corresponds to the second virial coefficient.

Here we employ the correlation of $B_s$ for \ce{CO_2} introduced by Holste \textit{et al.} \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0$ = 23.02991 cm$^3$/mol, $B_1$ = $-2.455297$ $\times$ 10$^3$ (Kcm$^3$/mol) and $B_2$ = -1.22675$\times$10$^7$ (K$^2$cm$^3$/mol).
At equilibrium, $P_s$ = $P$.
However, 
equilibrium pressure
We have found from preliminaries simulations that $\Delta G_{\text{sim}}$

 Actually, the prediction of $P_s$ should be iterative and in this work we propose the following scheme: (i) run an NPT simulation, (ii) calculate $\Delta G_{\text{sim}}$ and $f^{g}_s$ via Eq.~\ref{eq:fgas_d}, (iii) employ $f^{g}_s$ in Eq.~\ref{eq:fgas} to compute $P_s$ (iv) if $P_s$ $\neq$ P, repeat the above steps using $P_s$ as the specified pressure in a new NPT simulation. This is done until $P_s$ = P.

@@

The versatility of ILs for different applications is in part due to the fact that their physical properties strongly vary with the particular cation-anion combination. On the other hand, this characteristic also explains why the desirable transferability of force fields is particularly difficult to attain. Indeed, it is a common practice to extract the potential parameters corresponding to intramolecular (bond stretching, angle bending, torsion rotation) and Lennard-Jones (LJ) interactions from a well established force field and then derive the partial atomic charges. This procedure is, however, inaccurate as in conventional force fields the determination of the atomic charges precedes the dihedrals parametrization. Another ad-hoc strategy, which intends to crudely incorporate polarization effects, is to scale the partial charges uniformly with a factor which normally takes values in the range 0.8-0.9. This practice has proven to be effective in improving the predictions of dynamic properties \cite{Schr_der_2012}.

In this paper, we consider four models of \ce{[emim][B(CN)_4]}. Koller \textit{et al.} \cite{Koller_2012} were the first to introduce a specific non-polarizable force field of \ce{[emim][B(CN)_4]}. They tested three variations of a \textit{united-atoms} model, in which the alkyl groups \ce{CH_3} and \ce{CH_2} of \ce{[emim]^+} are replaced by pseudo-atoms. Among them, the ``FF-3'' model best reproduces the experimental values of density, self-diffusion coefficients and viscosity. In short, both the intramolecular and LJ parameters of the cation were taken from the work of Liu \textit{et al.} \cite{Liu_2006}, which in turn is based on the AMBER \cite{Cornell_1995} framework. The intramolecular parameters involving the boron atom were extracted from the DREIDING \cite{Mayo_1990} force field and the anionic vdW parameters were refined based on the values reported by Price \textit{et al.} \cite{Price_2001}. The atomic partial charges were derived using the EA-RESP method \cite{Basma_2001} from \textit{Post-Hartree-Fock ab-initio} calculations at MP2/6-31G$^\ast$+ level of theory, which automatically resulted in total charges on the ions of $\pm$0.8426$e$.

The potential parameters of \ce{[B(CN)_4]^-} derived by Koller \textit{et al.} \cite{Koller_2012} were then employed by Batista \textit{et al.} \cite{Batista_2015} to study aqueous solutions of \ce{[emim][B(CN)_4]}. In this case, the authors considered the \textit{all-atom} model of \ce{[emim]^+} developed by Cadena and Maginn \cite{Cadena_2006}. The atomic partial charges were obtained using the CHELP procedure \cite{Breneman_1990}, resulting in net charges of $\pm$0.889$e$ on the ions. It should be noted that those force fields consider different scaling factors for the 1-4 non-bonded interactions: Cadena and Maginn \cite{Cadena_2006} use values of 1.0 y 0.4 for the LJ and Coulombic interactions, respectively; while in Koller \textit{et al.} \cite{Koller_2012} the corresponding factors are 0.5 y 5/6, in accordance with the AMBER force field. Batista \textit{et al.} \cite{Batista_2015} employ the same factors as Cadena and Maginn \cite{Cadena_2006} (J.P. Coutinho, personal communication).

@@

We also consider the parametrization recently reported by Weber and Kirchner \cite{Weber_2016}, which is based on the force field developed by P{\'{a}}dua and Canongia Lopes~\cite{Canongia_Lopes_2006}. The potential parameters for the anion are not specified in Refs.~\citenum{Weber_2016} and~\citenum{Canongia_Lopes_2006}, and we obtained them by a direct communication with the authors of Ref.~\citenum{Weber_2016}. The atomic partial charges were derived using the RESP procedure from \textit{Hartree-Fock} calculations at 6-31++G$^{\ast \ast}$ level of theory, resulting in net charges of $\pm$0.8$e$.

Finally, for [emim][Tf\textsubscript{2}N] we use the model of K\"{o}ddermann~\textit{et al.} \cite{K_ddermann_2007}, which was derived by refining the force field due to P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}. This refining was characterized by their authors as ``unusual'', given that the LJ parameters were modified, while keeping the partial charges at their original values. This model was recently employed by Kerl \textit{et al.} \cite{Kerl__2017}, who were able to accurately predict Henry's constant of \ce{CO_2} at several temperatures.

@@

The solutes considered here are hexane, benzene, \ce{CO_2}, ethanol, cyclohexane and water. Hexane and benzene were modeled using the OPLS-AA force field \cite{Jorgensen_1996}, while the TRAPPE potential \cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}. Cyclohexane was simulated using a rigid-body model \cite{munoz2015lennard} whose sites only interact via the LJ potential and for water we employ the TIP4P/2005 model \cite{Abascal_2005,Vega_2011}.

@@

With the aim of setting up a (more) computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method \cite{Fennell2006} for computing the electrostatic interactions. As an alternative to the Ewald schemes \cite{Ewald_1921,Darden_1993,Hockney_1988}, McCann \textit{et al.} \cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force fields. They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ together with a cutoff distance ($r_c$) of 15 {\AA} provide the best energetic agreement with the Ewald method. In this work, the comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in good agreement for the ILs studied here. These values were also employed for the charged solutes, with the exception of hexane, for which we considered undamped Coulombic interactions. The reason for this exception is that we were able to improve the agreement with the experimental solvation free energy of that pure substance.

When it was possible to treat the ions as rigid bodies, the NPT algorithm of Kamberaj \textit{et al.} \cite{Kamberaj_2005} was employed to carry out the MD simulations. Otherwise, the scheme due to Shinoda \textit{et al}. \cite{Shinoda2004} was used along with the application of SHAKE \cite{Ryckaert1977} to constrain the C-H bonds lengths. For both integrators, we employed $\Delta t$ = 1 fs.

The gradual insertion/deletion of charged solutes involves scaling both the van der Waals (vdW) and the electrostatic intermolecular interactions. In order to avoid numerical instabilities associated to the vdW scaling, a standard approach is to employ a soft-core potential with a nonlinear coupling. In this paper we use the modified LJ potential due to Beutler \textit{et al.} \cite{Beutler_1994}. The electrostatic interactions, on the other hand, can be handled simply by a linear scaling of the potential \cite{Naden_2015}.

Furthermore, the contributions to $\Delta G$ originating at the different types of coupling can be taken into account separately, as follows
\begin{equation}
\begin{split}
\Delta G_{\text{sim}} =& \; \Delta G_{\,\text{vdW}} (\lambda_{\,\text{vdW}} = 0\rightarrow  1, \lambda_{\, \text{Coul}} = 0) \\
+ \; &\Delta G_{\,\text{Coul}} (\lambda_{\, \text{vdW}} = 1,\lambda_{\,\text{Coul}} = 0 \rightarrow 1),
\end{split}
\end{equation}
with $\lambda_{\,\text{vdW}}$ and $\lambda_{\,\text{Coul}}$ being the coupling parameters of the soft-core and electrostatic potentials, respectively. Note that in the process of coupling the vdW interactions, the Coulombic interactions are turned-off, while we set $\lambda_{\,\text{vdW}}$ = 1 when computing $\Delta G_{\,\text{Coul}}$.  This results in the ``end-point'' states given by ($\lambda_{\,\text{vdW}}$=0,$\lambda_{\,\text{Coul}}$=0) and ($\lambda_{\,\text{vdW}}$=1,$\lambda_{\,\text{Coul}}$=1), as expected. The complete sets of the coupling parameters we employ here are $\lambda_{\text{vdW}}$ = $\{0.0$, $0.05$, $0.1$, $0.2$, $0.25$, $0.3$, $0.35$, $0.4$, $0.45$, $0.5$, $0.6$, $0.7$, $0.8$, $0.9$, $1.0\}$ and $\lambda_{\text{Coul}}$ = $\{0.0$, $0.1$, $0.4$, $0.7$, $0.1\}$.

In this work, we carried out independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the states associated to the $\lambda_{\,\text{vdW}}$ set. In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$. This way, the sampling of the EE involves both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$. Thus, a remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation. In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT MD for a determined number of steps $N^{\text{EE}}$ followed by an attemp to change $\lambda_{\,\text{vdW}}$ using MC, and so on. That change in $\lambda_{\,\text{vdW}}$ is randomly proposed according to an independent sampling scheme \cite{Chodera_2011_2}. In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling. If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$ \cite{Lyubartsev_1992}. Of course, this can be achieved only iteratively. The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were all implemented in the LAMMPS software package \cite{Plimpton1995}, which we employed for carrying out all the MD simulations.

>>>>

%\documentclass[3p,preprint]{elsarticle}
\documentclass[3p,twocolumn]{elsarticle}
\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{commath}
\usepackage{amssymb}
\usepackage{scalerel}
\usepackage{natbib}

@@

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

@@

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

@@

Recently, the potentiality of ILs for \ce{CO_2} capture has been called into question by Carvalho \textit{et al.} \cite{Carvalho_2016}, who warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposed absorption capability of certain ILs. Although most thermodynamic models are formulated in terms of molar fraction, a key factor that dictates the absorption process efficiency is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent. By expressing the solubility values in terms of molality and performing an exhaustive comparison between various ILs, Ref. \cite{Carvalho_2016} shows that most systems which promote a physical absorption of \ce{CO_2} exhibit a solubility similar or even lower than that observed for the heavy alkane eicosane. This is true for the ``popular'' \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}; the only exception being the family of ILs containing the anion tetracyanoborate \ce{[B(CN)_4]^-}.

@@

In the following, the subscripts ``$s$'' and ``$\text{IL}$'' refer to the solute and the solvent (ionic liquid), respectively. For a system containing $N_{\text{IL}}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (P) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is formally defined as
\begin{equation}
\Delta G_{\text{sim}} \equiv k_b T \ln \frac{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 1)}{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 0)}, 
\end{equation}
where superscript ``$l$'' refers to the liquid solution, $\bigtriangleup$ denotes the isobaric-isothermal partition function, and $k_b$ is the Boltzmann's constant. Shirts \textit{et al.} \cite{Shirts_2003} rigorously derived the working expression that relates $\Delta G_{\text{sim}}$ to what is experimentally known as the solvation free energy. They found out that
\begin{equation}
\begin{split}
\label{eq:free_solv}
 \Delta G_{\text{solv}}& =  k_b T \ln \left( \frac{P_s}{c_s^{\ast} k_b T} \right)\\ &= \Delta G_{\text{sim}} - k_bT \ln \left( \frac{V^{\ast}}{V_1} \right) ,
\end{split}
\end{equation}
where $V^{\ast}$ and $V_1$ refer to the mean volumes of the systems specified by ($N_s^l-1$,$N_{\text{LI}}$,P,T) and ($N_s^l$,$N_{\text{LI}}$,P,T), respectively; $c_s^{\ast}$ = $N_s^l/V_1$ is the numerical concentration of the solute in the liquid phase and $P_s$ is the partial pressure of the solute in the gas phase. It should be noted that in the derivation of Eq.~\eqref{eq:free_solv}, the gas phase was treated as an ideal gas.

@@

\subsection*{Henry constant}
If both $P_s$ and $c_s^{\ast}$ were not large, we can resort to Eq.~\eqref{eq:free_solv} to predict Henry's law constant $\text{K}_s$, which is defined as \cite{Prausnitz}
\begin{equation}
\text{K}_s =\frac{P_s}{x_s}.
\end{equation}
Therefore, by simply expressing Eq.~\eqref{eq:free_solv} in terms of the molar fraction $x_s$ and assuming that MW (molecular weight of the solution) = $\text{MW}_{\scaleto{IL}{4pt}}$, as well as $\rho$ (mass density of the solution) = $\rho_{\scaleto{IL}{4pt}}$, we obtain that
\begin{equation}
\label{eq:henry_eq}
K_s = \tilde{\rho}_{\scaleto{IL}{4pt}} R \, T \exp(\beta \, \Delta G_{\text{sim}}),
\end{equation}
where $\tilde{\rho}_{\scaleto{IL}{4pt}}$ is the molar density of the solvent and $R$ is the gas constant.

@@

If one makes $N_s^{l}$ = 1, and substitute Eqs.~\eqref{eq:mu_res} and~\eqref{eq:mu_res_0} into Eq.~\eqref{eq:gamma}, the infinite-dilution activity coefficient, $\gamma^{\infty}_s$, can be predicted according to the following equation:
\begin{equation}
\label{eq:gamma_inf}
\ln \gamma^{\infty}_s = \frac{1}{k_b T} \left( \Delta G_{\text{sim}} - \Delta G_{\text{sim}}^{0} \right) + \frac{c_{\scaleto{LI}{4pt}}^{\, \ast}}{c_{\, s}^{\, \ast,0}},
\end{equation}
where $c_{\scaleto{IL}{4pt}}^{\, \ast}$ y $c_{\, s}^{\, \ast,0}$ are the numerical densities of the solvent and the pure solute, respectively.

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} are found at pressure values for which, in principle, the ideal gas assumption may be not valid. In this case, we shall correct the ideal gas contribution to the chemical potential, as follows
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + RT \ln \phi_s,
\end{equation}
where $\phi_s$ is the fugacity coefficient of the solute. This equation, together with the fact that $f^{g}_s$ = $P_s \phi_s$, are used to evaluate $\mu^{l}_s$ = $\mu^{g}_s$, which yields
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = \exp  \left( \beta \Delta G_{\text{sim}} \right) +  k_b T c_s^{\ast}, 
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase. Having calculated $f^{g}_s$, $P_s$ can be predicted using the following equation
\begin{equation}
\label{eq:fgas}
f^{g}_s = P_s \exp\left[ \frac{B_s(T) P_s}{R T} \right],
\end{equation}
where $B_s$ corresponds to the second virial coefficient. Here we employ the correlation of $B_s$ for \ce{CO_2} introduced by Holste \textit{et al.} \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0$ = 23.02991 cm$^3$/mol, $B_1$ = $-2.455297$ $\times$ 10$^3$ (Kcm$^3$/mol) and $B_2$ = -1.22675$\times$10$^7$ (K$^2$cm$^3$/mol).

As supported by preliminary simulations, it is reasonable to assume $\Delta G_{\text{sim}}$ as independent of pressure over the range of experimental data considered in this work. Then, the equilibrium pressure is determined by solving Eq.~\eqref{eq:fgas} for $P_s$.

@@

The versatility of ILs for different applications is in part due to the fact that their physical properties strongly depend on the particular cation-anion combination. On the other hand, this feature also explains why the desirable transferability of force fields is particularly difficult to attain. Indeed, it is a common practice to extract the potential parameters of the intramolecular (bond stretching, angle bending, torsion rotation) and Lennard-Jones (LJ) interactions from a well established force field and, then, only derive the partial atomic charges with the aim of ``adjusting'' the model for a specific pair of ions. This procedure is, however, inaccurate as in conventional force fields the determination of the atomic charges precedes the dihedrals parametrization. Another ad-hoc strategy, which intends to crudely incorporate polarization effects, is to scale the partial charges uniformly with a factor which normally takes values in the range 0.8-0.9. This has proven to be effective for predicting dynamic properties \cite{Schr_der_2012}.

In this paper, we consider four models of \ce{[emim][B(CN)_4]}. Koller \textit{et al.} \cite{Koller_2012} were the first to introduce a specific non-polarizable force field for that IL. They tested three variations of a \textit{united-atoms} model, in which the alkyl groups \ce{CH_3} and \ce{CH_2} of \ce{[emim]^+} are replaced by pseudo-atoms. Among them, the ``FF-3'' model best reproduces the experimental values of density, self-diffusion coefficients and viscosity. In short, both the intramolecular and LJ parameters of the cation were taken from the work of Liu \textit{et al.} \cite{Liu_2006}, which in turn is based on the AMBER \cite{Cornell_1995} framework. The intramolecular parameters involving the boron atom were extracted from the DREIDING \cite{Mayo_1990} force field and the anionic vdW parameters were refined based on the values reported by Price \textit{et al.} \cite{Price_2001}. The atomic partial charges were derived using the EA-RESP method \cite{Basma_2001} from \textit{Post-Hartree-Fock ab-initio} calculations at MP2/6-31G$^\ast$+ level of theory, which automatically resulted in total charges on the ions of $\pm$0.8426$e$.

The potential parameters of \ce{[B(CN)_4]^-} derived by Koller \textit{et al.} \cite{Koller_2012} were then used by Batista \textit{et al.} \cite{Batista_2015} to study aqueous solutions of \ce{[emim][B(CN)_4]}. In this case, the authors considered the \textit{all-atom} model of \ce{[emim]^+} developed by Cadena and Maginn \cite{Cadena_2006}. The atomic partial charges were obtained using the CHELP procedure \cite{Breneman_1990}, resulting in net charges of $\pm$0.889$e$ on the ions. It should be noted that the factors for scaling the 1-4 non-bonded interactions are different for these force fields: Cadena and Maginn \cite{Cadena_2006} apply factors of 1.0 and 0.4 for the LJ and Coulombic interactions, respectively; while in Koller \textit{et al.} \cite{Koller_2012} the corresponding values are 0.5 and 5/6, in accordance with the AMBER force field. Batista \textit{et al.} \cite{Batista_2015} used the first ones (J.P. Coutinho, personal communication).

@@

We also consider the parametrization reported by Weber and Kirchner \cite{Weber_2016}, which is based on the force field developed by P{\'{a}}dua and Canongia Lopes~\cite{Canongia_Lopes_2006}. The potential parameters for the anion are not specified in Refs.~\citenum{Weber_2016} and~\citenum{Canongia_Lopes_2006}, and we obtained them by a direct communication with the authors of Ref.~\citenum{Weber_2016}. The atomic partial charges were derived using the RESP procedure from \textit{Hartree-Fock} calculations at 6-31++G$^{\ast \ast}$ level of theory, resulting in net charges of $\pm$0.8$e$.

Finally, for [emim][Tf\textsubscript{2}N] we use the model of K\"{o}ddermann~\textit{et al.} \cite{K_ddermann_2007}, which was derived by refining the force field due to P{\'{a}}dua and Canongia Lopes \cite{Canongia_Lopes_2006}. This refining was characterized by their authors as ``unusual'', given that the LJ parameters were modified, while keeping the partial charges at their original values. This model was recently employed by Kerl \textit{et al.} \cite{Kerl__2017}, who were able to accurately predict Henry constants of \ce{CO_2} at several temperatures.

@@

The solutes considered in this work are hexane, benzene, \ce{CO_2}, ethanol, cyclohexane and water. Hexane and benzene were modeled using the OPLS-AA force field \cite{Jorgensen_1996}, while the TRAPPE potential \cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}. Cyclohexane was simulated using a rigid-body model \cite{munoz2015lennard} whose sites only interact via a LJ potential. For water, we employ the TIP4P/2005 model \cite{Abascal_2005,Vega_2011}.

@@

With the aim of setting up a computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method \cite{Fennell2006} for computing the electrostatic interactions. As an alternative to the Ewald schemes \cite{Ewald_1921,Darden_1993,Hockney_1988}, McCann \textit{et al.} \cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force fields. They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ and a cutoff distance ($r_c$) of 15 {\AA} provide the best energetic agreement with the Ewald method. In this work, the comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in a good agreement. These values were also employed for the charged solutes, with the exception of hexane, for which we applied undamped Coulombic interactions improving the prediction of the chemical potential.

The NPT algorithm of Kamberaj \textit{et al.} \cite{Kamberaj_2005} was used to perform the simulations with rigid bodies. For flexible molecules, we applied the scheme due to Shinoda \textit{et al.} \cite{Shinoda2004} C-H bonds were constrained by the SHAKE method \cite{Ryckaert1977}. For both integrators, we used a time-step size ($\Delta t$) of 1 fs.

The gradual insertion/deletion of charged solutes involves scaling both the van der Waals (vdW) and the electrostatic intermolecular interactions. In order to avoid numerical instabilities associated to the vdW scaling, a standard approach is to employ a soft-core potential with a nonlinear coupling. In this paper we use the modified LJ potential due to Beutler \textit{et al.} \cite{Beutler_1994}. The electrostatic interactions, on the other hand, can be handled by a linear scaling \cite{Naden_2015}. Furthermore, the contributions to $\Delta G$ originating at the different couplings are taken into account separately, as follows:
\begin{equation}
\begin{split}
\Delta G_{\text{sim}} =& \; \Delta G_{\,\text{vdW}} (\lambda_{\,\text{vdW}} = 0\rightarrow  1, \lambda_{\, \text{Coul}} = 0) \\
+ \; &\Delta G_{\,\text{Coul}} (\lambda_{\, \text{vdW}} = 1,\lambda_{\,\text{Coul}} = 0 \rightarrow 1),
\end{split}
\end{equation}
where $\lambda_{\,\text{vdW}}$ and $\lambda_{\,\text{Coul}}$ are the coupling parameters of the soft-core and electrostatic potentials, respectively. Note that in the process of coupling the vdW interactions, the Coulombic interactions are turned-off, while we set $\lambda_{\,\text{vdW}}$ = 1 when computing $\Delta G_{\,\text{Coul}}$.  This results in the ``end-point'' states given by ($\lambda_{\,\text{vdW}}$=0,$\lambda_{\,\text{Coul}}$=0) and ($\lambda_{\,\text{vdW}}$=1,$\lambda_{\,\text{Coul}}$=1), as expected. The complete sets of the coupling parameters we use here are $\lambda_{\text{vdW}}$ = $\{0.0$, $0.05$, $0.1$, $0.2$, $0.25$, $0.3$, $0.35$, $0.4$, $0.45$, $0.5$, $0.6$, $0.7$, $0.8$, $0.9$, $1.0\}$ and $\lambda_{\text{Coul}}$ = $\{0.0$, $0.1$, $0.4$, $0.7$, $0.1\}$.

In this work, we carried out independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the $\lambda_{\,\text{vdW}}$ set. In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$. This way, the sampling of an EE comprises both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$. Thus, a remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation. In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT molecular dynamics (MD) for a determined number of steps $N^{\text{EE}}$ followed by an attemp to change $\lambda_{\,\text{vdW}}$ using MC, and so on. This change in $\lambda_{\,\text{vdW}}$ is randomly proposed according to an independent sampling scheme \cite{Chodera_2011_2}. In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling. If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$ \cite{Lyubartsev_1992}. Of course, this can be achieved only iteratively. The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were implemented in the LAMMPS software package \cite{Plimpton1995}, which we used for carrying out all the simulations.

As demonstrated in Ref.~\citenum{Chodera_2011_2}, smaller values of $N^{EE}$ yield more decorrelated samples in each subensemble. In this work, we employ $N^{EE}$ = 5 which, together with the specified $\lambda_{\, \text{vdW}}$ set, resulted to be satisfactory as up to 3 preliminary runs of the EE were enough to determine the weights $\eta$ that yielded, in a final simulation, a ``sufficient'' sampling of each subensamble. By ``sufficient'', we mean that the sampling provides an acceptable uncertainty in $\Delta G$. The simulation times for the preliminary and final EE runs were 10 and 40 ns, respectively; and each $\lambda_{\text{Coul}}$ state was simulated for 10 ns. Finally, the MBAR method \cite{Shirts_2008}, as implemented by Klimovich \textit{et al.} \cite{Klimovich_2015}, was applied to estimate $\Delta G_{\text{vdW}}$ and $\Delta G_{\text{Coul}}$.
